A Multi-Modal Model of Object Deformation under Robotic Pushing

### 2016-12-25 twostagedo-rev-Dec-16.tex
merged changes from Vero into Jeremy’s changes

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
 F + F_{ext} = m \dfrac{\partial^2 p}{{\partial t}^2}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.

@@

This measurement \comment{is know as the F-score and} is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the score for the video \comment{ at frame $i$} is the mean over all the frames so far, $\mu(F)$:
\comment{\begin{align}
 \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
\end{align}}
If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

\section{\comment{Learning to predict forces}}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the \comment{same} training data set \comment{as the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.}

The \comment{stress strain curves} in [\fref{fig:sstrain}] are approximated by segmented regression curves.  Several candidates are proposed using the least squares technique and the best are chosen as follows.  Since there is only one independent variable the equations for adjusting a line are straightforward:
\begin{align}
 y & = mx + y_0 + Error \\
 m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
 y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
 Error & = \sum(y_i - (mx_i+y0))^2
\end{align}
Where $y$ is the dependent variable, $x$ the independent one, $m$ is the slope of the line, and $y_0$ the $y$ intercept.  To adjust logarithmic curves a change of variable is used.  Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.  In this case, the final equation will be of the form: $y = m \ln x + y_0$.  It would be possible to add other types of curves, but these were enough for the materials covered.

\comment{For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.  To determine the points of discontinuity between segments, each family is evaluated incrementally. The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.  We begin with a small number of points (e.g. five). To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.  This marks the point of discontiuity. Having discovered one best parameterized line for each family, the two resulting candidates are compared. We select the type of curve that covers more points in a single segment}.  This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.

\section{Experimental Results}
\label{sec:experiments}
\subsection{Scenario}
For the experimental scenario a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth, is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.  A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].  The force sensor was a DAQ-FT-Gamma, that can measure forces up to 200 Newtons in the $z$ direction (up to 65N in the others) and torques up to 5 Newton-metre.  The size of the photographs is $800 \times 600$ pixels.  The scale between the real world measurements and the pixels in the image were: $4600\frac{\unit{pixels}}{\unit{m}}$ \footnote{$\unit{m}=metre$} for the sponge and $4350\frac{\unit{pixels}}{\unit{m}}$ for the plasticine (they are different because the camera was placed at slightly different positions).

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Scores of Sets of Parameters Obtained With Different Integration Schemes}
\label{tab:ischemes}
\centering
\begin{tabular}{cccc}
\hline
\bfseries Scheme & \bfseries Best & \bfseries Worst & \bfseries Average \\
\hline\hline
Euler & 0.88 $\pm$ 0.04 & 0.44 $\pm$ 0.13 & 0.75 $\pm$ 0.08 \\
Verlet & 0.76 $\pm$ 0.19 & 0.53 $\pm$ 0.18 & 0.68 $\pm$ 0.19 \\
Customised & 0.921 $\pm$ 0.008 & 0.60 $\pm$ 0.13 & 0.86 $\pm$ 0.02 \\
\hline
\multicolumn{4}{p{88mm}}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.  \comment{Mean $\mu(F)$ and standard deviation across 24 training sesions are reported for the best performing models of every training session, worst model and average model.}}
\end{tabular}
\end{table}

The traditional equations of movement, estimated with Euler or Verlet show big oscillations and frequently become unstable.  Verlet seems to be badly affected by the collision resolution routines and geometric aids, that drastically affect the values of velocity and acceleration.  On the other hand, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that more closely resemble the ground truth, thus obtaining better marks.  From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them.

\subsection{\comment{Results of Automatic Parameter Calibration}}
\subsubsection{\comment{Performance for the Learned Shape Prediction Model}}
The performance of the genetic learning algorithm was evaluated using the customised integration scheme and the mass-spring elastic-plastic model.  At this point no prediction of the force was included, rather, the reaction forces measured with the force sensor were used.  \fref{fig:performance} shows how the quality of the simulation throughout the whole video improved with each generation.  Both graphs show that even though the enforcing of geometric constraints with additional routines may produce simulations with lower scores at the begining of the training, they perform just as well after enough generations.  These scores reflect the degree to which areas covered by the ground truth and the mesh of the simulation match; however, when looking at the videos produced, it becomes evident that simulations with additional geometric constraints succeed more in keeping the mesh flat. % Should put images? (if there is space left)

@@

\subsubsection{\comment{Performance for the Learned Force Prediction Model}}

@@

One hundred photographs and force readings were taken every $0.17\pm0.007 s$ in average.  The mass spring model runs with an integration step \comment{$h=0.1s$}, where forces that were not measured directly are interpolated.  The push down movement lasts approximately 50 frames and the retrieval movement the other 50.  The continuous lines in \fref{fig:sstrain} show the predicted forces, given the amount of compression of the material (strain), while the point-lines show the actual readings.  \fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the actual shapes predicted \comment{by the combined FP and SP model}.

@@

@@

The mass-spring models \comment{calibrated for the prediction task} were used to classify \comment{materials} through recognition of their behaviour, even \comment{when the objects where pushed and deformed at novel locations.}  The models make predictions as if the \comment{material of the} object was known.  The model that makes the best predictions indicates which \comment{type of} object is being observed.  There are two criteria for selecting an \comment{object/material for the video being shown to the system}:

\begin{enumerate}
 \item Per frame: \comment{a material is selected for each frame of the interaction according to whichever model made the best prediction for that particular frame, according to the F-score.} This test was performed for every frame of the six videos, where the robot finger pushes the block of material.  \tref{tab:classperframe} shows the results for the videos that were used as training and test sets for the prediction task.

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Classification per Frame}
\label{tab:classperframe}
\centering
\begin{tabular}{ccccccccc}
\hline
 & & \multicolumn{3}{c}{\bfseries Sponge} & & \multicolumn{3}{c}{\bfseries Plasticine} \\
Data Set & & TP & FN & Precision & & TP & FN & Precision\\
\hline\hline
 \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
\hline
   training & & 167 & 132 & 0.56 & & 186 & 113 & 0.62 \\
   test1 & & 167 & 132 & 0.56 & & 271 & 28 & 0.91 \\
   test2 & & 242 & 57 & 0.81 & & 289 & 10 & 0.97 \\ 
\hline
%\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
\end{tabular}
\end{table}

 \item Global: \comment{The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score.} The difference in the \comment{F-score} is added for all frames.  Frames where the distinction is more noticeable will receive a greater weight.
\end{enumerate}

\section{Discussion}
\label{sec:discussion}
We note the following points about the results and the method in general.
\subsection{The Learning/Calibration Method}
Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.  Better simulations have been obtained by hand tuning some of the parameters.  %\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.  The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}

\subsection{Using \comment{Additional} Geometric Constraints}
Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.   \comment{Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term}, even if its corresponding constant was small.  As a consequence the plastic material also recovered its original shape after some extra frames.  It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.  Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.  We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.  Meanwhile, the evaluation function gives the same weight to all frames in the sequence.  It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.  Future work will evaluate the impact of these modifications on the evaluation function.

@@

\subsection{Classification}
The best results for classification with the mass-spring model are obtained when the training of the models uses only forces to displace the nodes in the mass-spring model\footnote{Case 2 in \sref{sec:manysteps}}, but the combination of force readings and collision resolution is used during classification\footnote{Case 3 in \sref{sec:manysteps}}.  This finding agrees with the fact that the objects are more distinguishable from the data obtained from the force sensor, than from the visual data.  The fact that the spring constants must respond to the applied forces during the training and not only to the visual appearance is reflected in a better differentiation of the behaviour of the models.  When the calibration is more sensitive to the constants of the springs, it is more robust for novel interactions, therefore the difference in quality of the predictions is a better indicator of which material is being used.

>>>>

To simulate the effect of an external actuator (like the finger of the robot) interacting with the deformable object, external forces $F_{ext}$ acting on vertices of the mesh must be added to \eref{eq:newton}:
\begin{align}
F + F_{ext} = m \dfrac{\partial \vec{p}}{\partial t}
\end{align}
Another option could be to induce response forces by displacing vertices from their position of equilibrium.

@@

This measurement is known as the F-score and is used in the computer vision community to evaluate segmentation algorithms.  $F=1$ means that there were no false predictions ($FN+FP = 0$).  $F \rightarrow 0$\footnote{This is read: the value of $F$ goes to zero or $F$ tends to zero.}, if there are no good predictions ($TP\rightarrow0$) or if there are too many errors ($FN+FP \rightarrow \infty$).  $F$ is calculated for every frame and the score for the video at frame $i$ is the mean over all the frames so far, $\mu(F)$:
\begin{align}
 \mu(F) = \sum_{t=0}^{t=i} F(frame=t)
\end{align}
If the predictions for all the frames were perfect, $\mu(F)$ would be equal to 1; the closer to zero, the more errors there were.  The simulation stops sometimes, if a model has become unstable; from that frame on, the value of $F$ is zero.  The model with the highest final value of $\mu(F)$ is considered the best.

\section{Learning to predict forces}
%We used two methods to predict the forces.  The first one consists in measuring the total force acting on the vertex of the mesh closest to the actuator, according to the mass-spring model, after solving detected collisions.  This force is feed as input for the following simulation step.
\label{sec:forces}

Additionally and independently from the mass-spring system,  it is possible to learn to make predictions about the expected reaction force of the material, by using information from the stress-strain diagram obtained from the same training data set as the mass-spring model.  This diagram is obtained from the video recorded by the visual system of the robot and the forces recorded by a sensor located at the wrist.  When the readings of the force sensor start to increase and the figer collides with the block of material, the point of zero deformation is identified.  From here on, the displacement of the finger is considered equivalent to the amount of deformation in the material.  Even though this method is not as rigorous as an engineering test, the estimation is good enough for robot manipulation of common materials.

The stress strain curves in [\fref{fig:sstrain}] are approximated by segmented regression curves.  Several candidates are proposed using the least squares technique and the best are chosen as follows.  Since there is only one independent variable the equations for adjusting a line are straightforward:

\begin{align}
 y & = mx + y_0 + Error \\
 m & = \dfrac{n\sum x_i y_i-\sum x_i \sum y_i}{n \sum x_i^2 -(\sum x_i)^2} \\
 y_0 & = \dfrac{\sum y_i}{n} - y_0 \dfrac{\sum x_i}{n} \displaybreak[0] \\
 Error & = \sum(y_i - (mx_i+y0))^2
\end{align}
Where $y$ is the dependent variable, $x$ the independent one, $m$ is the slope of the line, and $y_0$ the $y$ intercept.  To adjust logarithmic curves a change of variable is used.  Before computing the line, the natural logarithm of the independent variable is calculated, that is: $x'_i = \ln(x_i)$.  In this case, the final equation will be of the form: $y = m \ln x + y_0$.  It would be possible to add other types of curves, but these were enough for the materials covered.

For automatic calibration of the force prediction model for the pushing (as opposed to retraction) phase both families of lines (logarithmic curves and straight lines) are considered.  To determine the points of discontinuity between segments, each family is evaluated incrementally. The candidate line for a particular family is obtained by fitting a growing subset of data points and calculating the fitting error.  We begin with a small number of points (e.g. five). To detect the discontinuity new points are added until the corresponding normalized error increases beyond a threshold value.  This marks the point of discontiuity. Having discovered one best parameterized line for each family, the two resulting candidates are compared. We select the type of curve that covers more points in a single segment.  This model is used to predict the forces for the mass-spring simulation \fref{fig:diagram} and \fref{fig:sstrain}.%The prediction can be averaged with the spring reaction force calculated from the previous step, to combine the visual and haptic modalities and test whether the quality of the prediction is improved.

\section{Experimental Results}
\label{sec:experiments}
\subsection{Scenario}
For the experimental scenario a rectangular block of deformable material, of $8.5 \unit{cm}$ long, $5 \unit{cm}$ wide and $1.8 \unit{cm}$ depth, is placed on a table; the robot pushes the object with a cylindrical finger, followed by a retrieval movement.  A colour firewire camera records the action perpendicularly, while a force sensor registers the reaction forces in synchrony with each photograph taken [\fref{fig:scene}(a)].  \comment{This set up allowed us to study the effect of using an automatically calibrated physics based model as model for the material, while using 2D meshes, even though the same principles can be used if we use a 3D point cloud and a 3D mesh.  To generate the initial mesh for the mass-spring model a routine placed the nodes for the masses evenly spaced on a rectangular grid, given the number of squares we wanted per side.  We made experiments with grids of $3x2, 10x5$ and $20x10$ squares.  To aliviate the bias produced by the springs they were placed in a star pattern, thus compensating forces on eight directions.}

The force sensor was a DAQ-FT-Gamma, that can measure forces up to 200 Newtons in the $z$ direction (up to 65N in the others) and torques up to 5 Newton-metre.  The size of the photographs is $800 \times 600$ pixels.  The scale between the real world measurements and the pixels in the image were: $4600\frac{\unit{pixels}}{\unit{m}}$ \footnote{$\unit{m}=metre$} for the sponge and $4350\frac{\unit{pixels}}{\unit{m}}$ for the plasticine (they are different because the camera was placed at slightly different positions).

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Scores of Sets of Parameters Obtained With Different Integration Schemes}
\label{tab:ischemes}
\centering
\begin{tabular}{cccc}
\hline
\bfseries Scheme & \bfseries Best & \bfseries Worst & \bfseries Average \\
\hline\hline
Euler & 0.88 $\pm$ 0.04 & 0.44 $\pm$ 0.13 & 0.75 $\pm$ 0.08 \\
Verlet & 0.76 $\pm$ 0.19 & 0.53 $\pm$ 0.18 & 0.68 $\pm$ 0.19 \\
Customised & 0.921 $\pm$ 0.008 & 0.60 $\pm$ 0.13 & 0.86 $\pm$ 0.02 \\
\hline
\multicolumn{4}{p{88mm}}{Where the mark is $\mu(F)$ as explained in \sref{sec:evaluation}.  Mean $\mu(F)$ and standard deviation across 24 training sesions are reported for the best performing models of every training session, worst model and average model.}
\end{tabular}
\end{table}

The traditional equations of movement, estimated with Euler or Verlet show big oscillations and frequently become unstable.  Verlet seems to be badly affected by the collision resolution routines and geometric aids, that drastically affect the values of velocity and acceleration.  On the other hand, the customised first degree equations suggested in \sref{sec:ischeme} produce smooth movements that more closely resemble the ground truth, thus obtaining better marks.  From this group of experiments it was also found that evolutionary learning is as efficient when considering all elastic and plastic terms, as when using only a few of them\footnote{\comment{The full list of terms is summarised at the end of \sref{sub:EvolAlg}.}}.

\subsection{Results of Automatic Parameter Calibration}
\subsubsection{Performance for the Learned Shape Prediction Model}
The performance of the genetic learning algorithm was evaluated using the customised integration scheme and the mass-spring elastic-plastic model.  At this point no prediction of the force was included, rather, the reaction forces measured with the force sensor were used.  \fref{fig:performance} shows how the quality of the simulation throughout the whole video improved with each generation.  Both graphs show that even though the enforcing of geometric constraints with additional routines may produce simulations with lower scores at the begining of the training, they perform just as well after enough generations.  These scores reflect the degree to which areas covered by the ground truth and the mesh of the simulation match; however, when looking at the videos produced, it becomes evident that simulations with additional geometric constraints succeed more in keeping the mesh flat. % Should put images? (if there is space left)

@@

\subsubsection{Performance for the Learned Force Prediction Model}

@@

One hundred photographs and force readings were taken every $0.17\pm0.007 s$ in average.  The mass spring model runs with an integration step $h=0.1s$, where forces that were not measured directly are interpolated.  The push down movement lasts approximately 50 frames and the retrieval movement the other 50.  The continuous lines in \fref{fig:sstrain} show the predicted forces, given the amount of compression of the material (strain), while the point-lines show the actual readings.  \fref{fig:simulation_sponge} and \fref{fig:simulation_plasticine} show the actual shapes predicted by the combined FP and SP model.

@@

\comment{We also compared the generalisation capabilities of our model against the system proposed in \cite{Cretu2012}.  This system uses a Growing Neural Gas (GNG) Network to calibrate a colour segmentation algorithm.  We couldn't use the HSV space as suggested because our videos have lots of shadows which produce holes inside the material regions which cause problems in further stages.  We got better results with Luv color space.  We clarify that, even though we used the HSV space in our tracking algorithm, we didn't have the same problems because the nodes of our linear snake search for nearby edges and are not affected by shadows in the middle of the object or temporary failures in detecting an edge.  Another GNG network is used to select the control points adequate for approximating the contour in the first frame.  Later, Cretu et al. use a Neural Gas (NG) algorithm to adjust that fixed number of control points to the contour of the material at every frame.  Using the coordinates of the finger, the sensed force and the coordinates of those control points through time, they train a Feedforward Neural Network (FFNN) to predict the contour of the material given the position of the finger and sensed force.  We reproduced those steps and their training method on our training video.  However, when we traid to test the generalisation capabilities of their method to new pushing actions, the weakness of the FFNN method became evident: it got tied to the specific location and type of interaction it was trained with, while our physics based model had no problem at making predictions for new pushing actions.  The results with this other method can be seen on \fref{fig:cretu_prediction}.

Since the videos with the plasticine have even greater shadows it was not posible to use their tracking algorithm on them.  However, even if we could change that algorithm, we would expect the system to have the same problems and predict the deformation at the middle of the material even if is being pushed somewhere else. }

\begin{figure}[!t]
\centering
\includegraphics[width=3.5in]{sponge__result.jpg}
\caption{Neural network approach fits quite well the training data set, but can not generalise to new pushing actions.  The green line shows the border used as ground truth, while blue lines show prediction from the FFNN.}\label{fig:cretu_prediction}
\end{figure}

@@

The mass-spring models calibrated for the prediction task were used to classify materials through recognition of their behaviour, even when the objects where pushed and deformed at novel locations.  The models make predictions as if the material of the object was known.  The model that makes the best predictions indicates which type of object is being observed.  There are two criteria for selecting an object/material for the video being shown to the system:

\begin{enumerate}
 \item Per frame: a material is selected for each frame of the interaction according to whichever model made the best prediction for that particular frame, according to the F-score. This test was performed for every frame of the six videos, where the robot finger pushes the block of material.

 \item Global: The system makes a vote about which material the object was made of, for the whole duration of the video, by making use of a weighted score. The difference in the F-score is added for all frames.  Frames where the distinction is more noticeable will receive a greater weight.
\end{enumerate}

\section{Discussion}
\label{sec:discussion}
We note the following points about the results and the method in general.
\subsection{The Learning Method}
Even though the evolutionary algorithm allowed for a uniform exploration of the logarithmic space of parameters, it failed to find the optimal solution.  Better simulations have been obtained by hand tuning some of the parameters.  %\sout{A factor that this learning algorithm fails to take into account are the significant differences in behaviour when the magnitude of a parameter is modified in factors of 10.  The random uniform distribution used to select parameters favors a given order of magnitude while the system would benefit more from stronger variantions here.}

\subsection{Using Additional Geometric Constraints}
Adding extra routines to enforce the planarity of the mesh had good aesthetic results, however that did not reflect much on the mark for the best videos of each generation during training.   Also, when attempting to model the plastic material, the geometric constraints provoked a much larger effect caused by the plasticity term, even if its corresponding constant was small.  As a consequence the plastic material also recovered its original shape after some extra frames.  It has been seen that, if the constant for linear preservation is set to zero, the problem disappears.  Even if the automatic calibration selects these extreme values with probability $\epsilon>0$, in an attempt to reproduce this finding, their derivatives obtained after applying gaussian noise to their parameters obtain better marks, even though they eventually recover their shapes, and get selected in a better position for the next generation.  We hypothesise that humans give greater weight to specific aspects of the simulation, for example: to the beginning, to the instant of greatest deformation and to the final shape.  Meanwhile, the evaluation function gives the same weight to all frames in the sequence.  It is possible to use the force learning algorithm and the commands given to the robot (push/retrieve) to detect these points, since they are correlated with discontinuities in these other modalities.  Future work will evaluate the impact of these modifications on the evaluation function.

@@

\begin{table}[!t]
\renewcommand{\arraystretch}{1.3}
\caption{Classification per Frame}
\label{tab:classperframe}
\centering
\begin{tabular}{ccccccccc}
\hline
 & & \multicolumn{3}{c}{\bfseries Sponge} & & \multicolumn{3}{c}{\bfseries Plasticine} \\
Data Set & & TP & FN & Precision & & TP & FN & Precision\\
\hline\hline
 \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
\hline
   training & & 167 & 132 & 0.56 & & 186 & 113 & 0.62 \\
   test1 & & 167 & 132 & 0.56 & & 271 & 28 & 0.91 \\
   test2 & & 242 & 57 & 0.81 & & 289 & 10 & 0.97 \\ 
\hline
\multicolumn{9}{p{0.9\columnwidth}}{\begin{footnotesize}\comment{* TP and FN measured in number of frames. TP + FN = total number of frames.  Final class assigned according to the performance of the models' predictions across the whole video.} \end{footnotesize}}
\end{tabular}
\end{table}
% \begin{table}[!t]
% \renewcommand{\arraystretch}{1.3}
% \caption{Classification per Frame}
% \label{tab:classperframe}
% \centering
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Sponge} \\
% Data Set & & TP & FN & TN & FP & Precision \\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 167 & 132 & & & 0.56  \\
%    test1 & & 167 & 132 & & & 0.56  \\
%    test2 & & 242 & 57 & & & 0.81 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% 
% \vspace*{1em}
% \begin{tabular}{ccccccccc}
% \hline
%  & & \multicolumn{4}{c}{\bfseries Plasticine} \\
% Data Set & & TP & FN & TN & FP & Precision\\
% \hline\hline
%  \multicolumn{9}{c}{Trained with forces on vertices and no collision resolution.} \\
% \hline
%    training & & 186 & 113 & & & 0.62 \\
%    test1 & & 271 & 28 & & & 0.91 \\
%    test2 & & 289 & 10 & & & 0.97 \\ 
% \hline
% %\multicolumn{7}{p{68mm}}{\begin{footnotesize}* TP and FN measured in number of frames.  Final class assigned according to the performance of the models' predictions across the whole video. \end{footnotesize}}
% \end{tabular}
% \end{table}

\comment{
\tref{tab:classperframe} shows the results for the classification task, frame per frame, for the videos that were used as training and test sets for the prediction task.  When the sponge model performed better for a frame of the sponge video it is marked as a true positive, when plasticine does better it is marked as a false negative (since it didn't recognize it as a sponge).  The same process is used for the other material.  Since the concept was proved with only two materials, true positives and false negatives provide all the information.

It calls for attention that in some of the experiments, like the one illustrated above, the performance of classification for the sponge on the training and test1 sets are the same.  We interpret that the mass-spring physics based model generalizes naturally when the force/deformation conditions are similar, even if the place for pushing is different.  Between pushing in the middle and pushing closer to the border the qualitative behaviour of the elastic material was no so different for these models; unlike the data from the third set, where the distribution of mass and maximum force/deformation is more different from the ones seen on the training set of the model.  It is interesting to confirm that the difference in predictions between the models is bigger here in favour of a better classification.

It is also remarkable that, even though, when modeling the sponge, the parameters for the plasticine were close in performance to the parameters for sponge (thus causing more problems to establish a good classification), the opposite does not happen.  That is, the parameters for the sponge are not good at modeling the plasticine on the test sets, thus yielding better results at the classification task.  This could happen because the plastic behaviour (that is, the permanent deformation of a spring) gets activated with stronger forces after the $yield$ limit is reached, which only appear when pushing the plasticine.  With the small forces feedback by the sponge, the model for the plasticine does not reach its range of plastic behaviour and thus the big differences in behaviour do not manifest.
}

The best results for classification with the mass-spring model are obtained when the training of the models uses only forces to displace the nodes in the mass-spring model\footnote{Case 2 in \sref{sec:manysteps}}, but the combination of force readings and collision resolution is used during classification\footnote{Case 3 in \sref{sec:manysteps}}.  This finding agrees with the fact that the objects are more distinguishable from the data obtained from the force sensor, than from the visual data.  The fact that the spring constants must respond to the applied forces during the training and not only to the visual appearance is reflected in a better differentiation of the behaviour of the models.  When the calibration is more sensitive to the constants of the springs, it is more robust for novel interactions, therefore the difference in quality of the predictions is a better indicator of which material is being used.
